---
title: "Data Science Homework #3"
author: "Sophia Miller"
date: "10/6/2019"
output: github_document
---
# Initial Set Up

```{r setupcode, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1
## Loading Instacart data

```{r load_instacart}
library(p8105.datasets)
data("instacart")
```

## Description of Instacart Dataset

The dataset used here is a cleaned and limited subset of the original Instacart data. Each row in this dataset represents a single product from an Instacart order. There are `r nrow(instacart)` rows and `r ncol(instacart)` columns in the dataset. Each observation has an order identifier (`order_id`), product identifier (`product_id`), and customer identifier (`user_id`). Key variables include `reordered`, which indicates  if the product has been ordered by the customer in the past month, `order_dow`, which indicates which day of the week the order was placed, `days_since_prior_order`, which indicates the number of days since the customer's last order, and `product_name`, which is the name of the product ordered. For example, the mean number of days between orders is `r mean(pull(instacart, days_since_prior_order))`.

## Exploratory data analysis of Instacart data

 
```{r aisles}
instacart %>% 
 summarize(
   n_aisles = n_distinct(aisle_id))

## change this to inline r code?
```

There are 134 aisles in the Instacart dataset.

```{r top_aisles}
names(which.max(table(pull(instacart, aisle))))

## return top three aisles rather than top one
```
The most items are ordered from the Fresh Vegetables aisle.

```{r aisle_plot}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_items_ordered = n()
  ) %>% 
   
  filter(n_items_ordered > 10000) %>% 
  ggplot(aes(x = n_items_ordered, y = aisle )) + geom_point()
```


```{r popular_items}
instacart %>% 
  group_by(aisle) %>%
    summarize(
    n_items_ordered = n())

## come back to this 
```





