---
title: "Data Science Homework #3"
author: "Sophia Miller"
date: "10/6/2019"
output: github_document
---
# Initial Set Up

```{r setupcode, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1
## Loading Instacart data

```{r load_instacart}
library(p8105.datasets)
data("instacart")
```

## Description of Instacart Dataset

The dataset used here is a cleaned and limited subset of the original Instacart data. Each row in this dataset represents a single product from an Instacart order. There are `r nrow(instacart)` rows and `r ncol(instacart)` columns in the dataset. Each observation has an order identifier (`order_id`), product identifier (`product_id`), and customer identifier (`user_id`). Key variables include `reordered`, which indicates  if the product has been ordered by the customer in the past month, `order_dow`, which indicates which day of the week the order was placed, `days_since_prior_order`, which indicates the number of days since the customer's last order, and `product_name`, which is the name of the product ordered. For example, the mean number of days between orders is `r mean(pull(instacart, days_since_prior_order))`.

## Exploratory data analysis of Instacart data

```{r aisles}
instacart %>% 
 summarize(
   n_aisles = n_distinct(aisle_id))

## change this to inline r code?
```

There are 134 aisles in the Instacart dataset.

```{r top_aisles}
names(which.max(table(pull(instacart, aisle))))

## return top three aisles rather than top one
```
The most items are ordered from the Fresh Vegetables aisle.

## Plot aisle vs. number of items ordered

```{r aisle_plot}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_items_ordered = n()) %>% 
  mutate(aisle = reorder(aisle, n_items_ordered)) %>% 
  filter(n_items_ordered > 10000) %>% 
  ggplot(aes(x = aisle, y = n_items_ordered)) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + geom_col() + ggtitle("Figure1. Number of items ordered per aisle") + theme(plot.title = element_text(hjust = 0.5))
```


## Make a table showing the three most popular items in "baking ingredients", "dog food care", and "packaged fruits and vegetables"

```{r popular_items}
instacart %>% 
  group_by(product_name, aisle) %>%
  summarize(
    n_items_ordered = n()
    ) %>% 
  group_by(aisle) %>% 
  filter(
    aisle %in% c("baking ingredients","dog food care", "packaged vegetables fruits"),
    min_rank(desc(n_items_ordered)) < 4) %>% 
  arrange(n_items_ordered, aisle) %>% 
  knitr::kable(format = 'pandoc', caption = "Table 1. Most popular items by aisle")
#check
```


## Make a table showing mean time of day pink lady apples and coffee ice cream are ordered on each day of the week

```{r time_of_day_table}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  mutate(order_dow = as.factor(order_dow)) %>% 
  arrange(order_dow) %>% 
  mutate(Day = recode(order_dow, '0' = "Sunday", '1' = "Monday", '2' = "Tuesday", '3' = "Wednesday", '4' = "Thursday", '5' = "Friday", '6' = "Saturday")) %>% 
  group_by(product_name, Day) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = product_name,
    values_from = mean_hour) %>% 
  knitr::kable(format = 'pandoc', caption = "Table 1. Mean hour of day of item order, by day of week", digits = 0)
```

# Problem 2






